---
title: "DATa606_Chapter 8 - Multiple and Logistic Regression"
author: "Yun Mai"
date: "May 4, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Chapter 8 - Multiple and Logistic Regression

```
Practice: 8.1, 8.3, 8.7, 8.15, 8.17
```


8.1 Baby weights, Part I. The Child Health and Development Studies investigate a range of topics. One study considered all pregnancies between 1960 and 1967 among women in the Kaiser Foundation Health Plan in the San Francisco East Bay area. Here, we study the relationship between smoking and weight of the baby. The variable smoke is coded 1 if the mother is a smoker, and 0 if not. The summary table below shows the results of a linear regression model for predicting the average birth weight of babies, measured in ounces, based on the smoking status of
the mother.

```{r,echo=F}
table8.1 <- data.frame(v1 = c(123.05, -8.94),v2 =c(0.65,1.03),v3 =c(189.60,-8.65),v4= c(0.0000,0.0000))
colnames(table8.1)<-c("Estimate","Std. Error","t value","Pr(>|t|)")
row.names(table8.1)<-c("(Intercept)","smoke")
library(knitr)
kable(table8.1)
```


The variability within the smokers and non-smokers are about equal and the distributions are symmetric. With these conditions satisfied, it is reasonable to apply the model. (Note that we don't need to check linearity since the predictor has only two levels.) 

(a) Write the equation of the regression line.


$$\hat{bodyweight} = 123.05 - 8.94 \times smoke$$

(b) Interpret the slope in this context, and calculate the predicted birth weight of babies born to smoker and non-smoker mothers.

The slope here is the difference of predicted body weight of babies born to smoker and non-smoker mothers.

the predicted birth weight of babies born to smoker: $$\hat{bodyweight}[smoker] = 123.05 - 8.94 \times 1 = 114.11$$

the predicted birth weight of babies born to non-smoker: $$\hat{bodyweight}[smoker] = 123.05 - 8.94 \times 0 = 123.05$$


(c) Is there a statistically significant relationship between the average birth weight and smoking?

The hypotheses:
H0: The the difference of estimated body weight of babies born to smoker and non-smoker mothers is zero.
HA: The the difference of estimated body weight of babies born to smoker and non-smoker mothers is zero.

The p-value corresponds exactly to the two-sided test we are interested in: 0. The p-value is so small
that we reject the null hypothesis and conclude that body weight of babies and smoking are negatively correlated.

8.3 Baby weights, Part III. We considered the variables smoke and parity, one at a time, in modeling birth weights of babies in Exercises 8.1 and 8.2. A more realistic approach to modeling infant weights is to consider all possibly related variables at once. Other variables of interest
include length of pregnancy in days (gestation), mother's age in years (age), mother's height in inches (height), and mother's pregnancy weight in pounds (weight). Below are three observations from this data set.
!()

